The Concurrent Calculi Formalisation Benchmark

### 2023-10-10 paper/paper.tex
Revise introduction + add package "fixme"

\section{Design of the benchmark}\label{sec:design-discussion}
\subsection{General}
- Why are the challenges not covered by POPLMark/POPLMark Reloaded?
- Choice in general
  - Choice can be encoded, so is not necessary
  - Complicates the theory
  - Add as ``future work''
- Extension categories: other topics and small technical differences
- Why not unified syntax?
  - Choice and coinduction does not mix
- A goal is to build a culture of formalising things in the community
- Encoding techniques are not different from any other field, so there's nothing ``special'' about the formalisations here

\subsection{Linearity}
- Linearity type system dual vs. Vasco's?
- Only linear channels or also unrestricted channels?
- Pi-calculus or CCS for linearity?
- Linearity: the context is polymorphic
- Linearity is generally interesting
- Why explicitly bind two endpoints in linearity? To make the duality/polarity connection obvious
- Linearity: Do we care about avoiding deadlocks? No, this is stronger than safety as in well-formedness
- Asynchrony? Not necessary, this is a it more advanced

\subsection{Scope extrusion}
- Choice in scope extrusion
- Scope extrusion does not behave as ``normal'' a binder wrt. semantics, so the usual techniques may not work
- Why do we keep the restrictions around in the LTS semantics in scope extrusion?

\subsection{Coinduction}
- Different notions of bisimulation in the coinduction challenge
- Coinduction: why is congruence interesting?
- We don't care about the content of messages in coinduction, only synchronization
- ``Proper'' barbs with a structural congruence in coinduction instead of LTS semantics

@@

As with the challenges that came before, this one is intended to be
useful on its own, but not to cover every problem that may be faced
when mechanising systems based on message passing calculi. The current
challenge can be extended in two dimensions: the first, about taking
the existing challenges further. And the second, about devising
challenges not covered by the current benchmarks.

The proposed challenges try to minimise the overlap of the techniques
they exercise, however, most systems will combine some or all aspects
of the challenges. An interesting second round of formalisations could
entail combining the features of the challenges. In the following
table we consider the following three additions: First the addition of
choice to the type level and the process level. Second, the addition
of recursion (and recursion types for typed systems). And finally, the
addition of channel delegation to those calculi that do not already
support it.

The following table shows which extension should be suitable (that we
mark with $+$) for each challenge:

@@

\begin{center}\small
  \begin{tabular}{|r|c|c|c|}
    \hline
    & Linearity & \shortstack{Scope \\ extrusion} & Coinduction \\
    \hline
    Choice & $+$ & $+$  & $+$ \\
    \hline
    \shortstack{Recursive \\ types} & $+$  & $+$ & \\
    \hline
    Delegation & $+$ & & $+$ \\
    \hline
  \end{tabular}
\end{center}

@@

The other dimension to extend these calculi corresponds to systems
that are not covered by our challenges. Extending in this dimension
means proposing new challenges to cover for different features of
message passing calculi, or to cover different features of mechanised
proof.

Some interesting aspects of message passing calculi to explore in
further challenges could be: Multiparty session types, and
choreographies because their definitions and meta-theory have elements
that we do not cover here, however the elements that we do cover will
appear in these problems. Also, while we explore a form of barbed
bisimilarity one could easily design a challenge exploring different
notions of bisimilarity (barbed, or weak, etc.), and trace
equivalence.

On the side of proof assistants, the current challenge proposes
mechanising the proofs of theorems. An interesting venue to explore
would be to take advantage of other aspects of proof assistants. For
example, proof assistants are often able to produce certified code (by
using code extraction, or compiling and running their definitions).
Additionally, while having effective automation in our current
challenge is possible, a challenge could be designed with the explicit
objective of automating aspects of the proofs. And finally, a
challenge could propose the integration with other formal reasoning
tools, namely model checkers. These tools are in common use in the
field. However, there combining automated proofs with proof assistants
could be a great path for attempting larger proofs with less effort.

Ultimately, the current challenge can be extended in several
worthwhile directions, and we look forward to a future when they
indeed are extended. Moreover, we see the current challenge as setting
the foundation for those future extensions. It is our expectation that
the solutions of those hypothetical challenges will build on solutions
to this challenge.

>>>>

\section{Design of the benchmark}\label{sec:design-discussion}
\subsection{Topics we discussed during the meetings}
- Linearity type system dual vs. Vasco's?
- Only linear channels or also unrestricted channels?
- Pi-calculus or CCS for linearity?
- Choice in general
  - Choice can be encoded, so is not necessary
  - Complicates the theory
  - Add as ``future work''
- Choice in scope extrusion
- Extension categories: other topics and small technical differences
- Why not unified syntax?
  - Choice and coinduction does not mix
- Different notions of bisimulation in the coinduction challenge
- Linearity: the context is polymorphic
- Coinduction: why is congruence interesting?
- Why are the challenges not covered by POPLMark/POPLMark Reloaded?
- Scope extrusion does not behave as ``normal'' a binder wrt. semantics, so the usual techniques may not work
- A goal is to build a culture of formalising things in the community
- Linearity is generally interesting
- Encoding techniques are not different from any other field, so there's nothing ``special'' about the formalisations here
- We don't care about the content of messages in coinduction, only synchronization
- Why explicitly bind two endpoints in linearity? To make the duality/polarity connection obvious
- Linearity: Do we care about avoiding deadlocks? No, this is stronger than safety as in well-formedness
- Asynchrony? Not necessary, this is a it more advanced
- Why do we keep the restrictions around in the LTS semantics in scope extrusion?
- ``Proper'' barbs with a structural congruence in coinduction instead of LTS semantics

@@

As with the challenges that came before, this challenge is intended to
have useful solutions, but not to cover every problem that may be
faced when mechanised systems based on message passing calculi. Going
beyond the current challenges could be in one of two dimensions: the
first dimension about making harder challenges starting from the
current ones, and a second one about devising challenges not covered
by the current benchmarks.

The proposed challenges are orthogonal, however, in general systems
will combine some or all aspects of the challenges. An interesting
second round of formalisations could entail combining the features of
the challenges. In the following table we consider the following three
additions: First the addition of choice to the type level and the
process level. Second, the addition of recursion (and recursion types
for typed systems). And finally, the addition of channel delegation
for those calculi that do not already support it.

The following table shows which extension should be suitable (denoted
by a $+$) for each challenge:

@@

\begin{tabular}{|r|c|c|c|}
  \hline
  & Linearity & \shortstack{Scope \\ extrusion} & Coinduction \\
  \hline
  Choice & $+$ & $+$  & $+$ \\
  \hline
  \shortstack{Recursive \\ types} & $+$  & $+$ & \\
  \hline
  Delegation & $+$ & & $+$ \\
  \hline
\end{tabular}

@@

Note that these additions, make the challenges less orthogonal by
combining some of their features. This is to be expected, in the sense
that the proposed challenges should form the building blocks of more
complex systems.

The other dimension to extend these calculi corresponds to systems
that are not covered by our challenges\dots
% Multiparty session types, and choreographies.

% different notions of bisimilarity (barbed, or weak, etc.), trace equivalence.

% Exercising different features of proof assistants like: code
% extraction and/or code automation.

% Interaction between formalised proofs and model checkers.

% And any other
